*Supplementary Material 1*

### 2022-06-23 ms/si.Rmd
Update SI figures
- fix spacing in trees
- don't plot Zygnema

```{r seqlen, fig.width = 7, fig.height = 4, eval = params$doc_type != "manual"}
fern_slen_tib <- read_from_cache(
  "fern_slen_tib", here::here("_supp_store")) %>%
  # Add column for plastome or not
  mutate(
    plastome = if_else(
      str_detect(def, "complete genome"),
      "A. 'complete genome'",
      "B. other"
    ),
    plastome = factor(plastome, levels = c("A. 'complete genome'", "B. other"))
  )

@@

```{r fossil-map-make}
# Make tibble with node number and label for ggtree
node_labs_tbl <-
  con_fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(
      taxon_1, taxon_2,
      ~ape::getMRCA(sanger_con_tree_rooted_pruned, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
fossil_map_plot <-
  ggtree(
    sanger_con_tree_rooted_pruned,
    branch.length = "none", size = 0.1, color = "grey30") %<+% node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 0.6) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"),
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(
    aes(label = fossil_label), hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r fossil-map-save, eval = params$doc_type == "doc"}
ggsave(
  plot = fossil_map_plot,
  file = (result_file(s_figure("fossil_map"), "pdf")),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("fossil_map")`**.
Consensus phylogeny (Sanger dataset) with node labels indicating placement of fossil constraints (`r result_file(s_figure("fossil_map"), "pdf") %>% fs::path_file()`).
Number for each constraint is age in Ma.

@@

```{r locus-upset, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Make tibble with one row per locus per species
sanger_alignment_tbl %>%
  mutate(taxon = map(align_trimmed, rownames)) %>%
  select(locus = target, species = taxon) %>%
  unnest(cols = species) %>%
  # Exclude outgroups
  left_join(
    unique(select(sanger_sampling, species, outgroup)), by = "species") %>%
  filter(outgroup == FALSE) %>%
  select(-outgroup) %>%
  group_by(species) %>%
  summarize(locus = list(locus)) %>%
  ggplot(aes(x = locus)) +
  geom_bar() +
  labs(x = "Locus combination", y = "Count (num. species)") +
  scale_x_upset(n_intersections = 20)
```

@@

```{r bs-plot}
phy <- sanger_con_tree_rooted

@@

# Make plot
bs_plot <- 
  ggtree(
    phy,
    size = 0.1, color = "grey30") %<+% bs_tibble +
  # Add node points
  geom_nodepoint(
    aes(color = bs), size = 0.1) +
  scale_color_viridis_c(
    "BS",
    option = "B", na.value = "transparent",
    labels = scales::percent
    ) +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r bs-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bs_plot,
  file = result_file(s_figure("bs"), "pdf"),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("bs")`**.
Consensus phylogeny (Sanger dataset) with node labels showing bootstrap support (%) (`r result_file(s_figure("bs"), "pdf") %>% fs::path_file()`).
Darker colors indicate lower support.
The backbone (plastome) phylogeny was used as a constraint tree, so all clades present in the backbone phylogeny tree receive 100% bootstrap support.

@@

```{r backbone-full-plot}
phy <- plastome_tree_rooted

@@

# Make plot
bb_full_plot <-
  ggtree(phy, size = 0.2, color = "grey10") %<+% 
  bs_tibble +
  # now we can show BS values using geom_text()
  geom_text(
    aes(label = bs), hjust = -.25, size = 1.0,
    color = "grey50") +
  # add tip labels
  geom_tiplab(aes(label = label), size = 1.0)
```

```{r backbone-full-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bb_full_plot,
  file = (result_file(s_figure("backbone_full"), "pdf")),
  height = 20, width = 10, units = "in",
  limitsize = FALSE)
```

**`r s_figure("backbone_full")`**.
Consensus plastome (backbone) phylogeny (`r result_file(s_figure("backbone_full"), "pdf") %>% fs::path_file()`).
Values at nodes are bootstrap support (%); only values less than 100% are shown.

@@

```{r du-ages, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Helper function for subsetting dates to get into
# almost-long format
subset_ages <- function(other_dates, author_sel) {
  other_dates %>%
    select(family, contains(glue("{author_sel}_"))) %>%
    rename_with(~str_remove_all(., glue("{author_sel}_"))) %>%
    mutate(author = author_sel)
}

@@

```{r comp-fossils, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
ts_fossil_calibration_tips %>%
  select(minimum_age, node_calibrated) %>%
  mutate(dataset = "ts") %>%
  bind_rows(
    con_fossil_calibration_tips %>%
      select(minimum_age, node_calibrated) %>%
      mutate(dataset = "current")
  ) %>%
  mutate(node_calibrated = fct_reorder(node_calibrated, minimum_age)) %>%
  ggplot(aes(x = minimum_age, y = node_calibrated, fill = dataset)) +
  geom_col(position = position_dodge()) +
  scale_x_continuous("Age (millions of years)", expand = expansion(mult = c(0,0.1))) +
  theme_bw() +
  scale_fill_manual(
    values = c(current = "grey30", ts = "grey80"),
    labels = c(current = "This Study", ts = "Testo and Sundue 2016")
  ) +
  theme(
    axis.title.y = element_blank(),
    legend.title = element_blank(),
    axis.text.y = element_text(size = 6),
    legend.position = "bottom"
  )
```

@@

```{r comp-stem-ages, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
# Combine family stem ages into almost-long format:
# one row per family/study combination,
# cols for median, high, and low date
all_family_stem_ages <-
  bind_rows(
    subset_ages(other_dates, "ts"),
    transmute(
      ts_family_stem_ages,
      family,
      author = "current",
      median = age
    )
  ) %>%
  left_join(
    unique(select(sanger_sampling, family, major_clade)), by = "family") %>%
  left_join(clade_tibble, by = "major_clade") %>%
  # (some median dates are NA if that family wasn't included in a study)
  assert(not_na, family, major_clade, code, author) %>%
  # Set family as a factor ordered by current age estimate
  left_join(ts_family_stem_ages, by = "family") %>%
  mutate(family = fct_reorder(family, age)) %>%
  select(-age)

@@

```{r ts-fossil-map-make}
# Make tibble with node number and label for ggtree
ts_node_labs_tbl <-
  ts_fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(
      taxon_1, taxon_2,
      ~ape::getMRCA(sanger_con_tree_rooted_pruned, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
fossil_map_plot <-
  ggtree(
    sanger_con_tree_rooted_pruned,
    branch.length = "none",
    size = 0.1, color = "grey30") %<+% ts_node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 0.6) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"),
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(
    aes(label = fossil_label), hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r ts-fossil-map-save, eval = params$doc_type == "doc"}
ggsave(
  plot = fossil_map_plot,
  file = (result_file(s_figure("ts_fossil_map"), "pdf")),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("ts_fossil_map")`**.
Consensus phylogeny (Sanger dataset) with node labels indicating placement of fossil constraints of @Testo2016a.
Number for each constraint is age in Ma.

@@

```{r comp-stem-points, fig.width = 6, fig.height = 5, eval = params$doc_type != "manual"}
# For checking that model data is same as plot data
mod_fitted_data <-
ftol_ts_comp_mod %>%
  broom::augment() %>%
  arrange(ftol)

@@

```{r ltt, fig.width = 7, fig.height = 6, eval = params$doc_type != "manual"}
# Make lineages through time (LTT) plot comparing Angiosperms and polypods

>>>>

```{r seqlen-plot, fig.width = 7, fig.height = 4, eval = params$doc_type != "manual"}
fern_slen_tib <- read_from_cache(
  "fern_slen_tib", here::here("_supp_store")) %>%
  # Add column for plastome or not
  mutate(
    plastome = if_else(
      str_detect(def, "complete genome"),
      "A. 'complete genome'",
      "B. other"
    ),
    plastome = factor(plastome, levels = c("A. 'complete genome'", "B. other"))
  )

@@

```{r fossil-map-plot-make}
# Make tibble with node number and label for ggtree
node_labs_tbl <-
  con_fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(
      taxon_1, taxon_2,
      ~ape::getMRCA(sanger_con_tree_rooted_pruned, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
fossil_map_plot <-
  ggtree(
    sanger_con_tree_rooted_pruned,
    branch.length = "none", size = 0.1, color = "grey30") %<+% node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 0.6) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"),
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(
    aes(label = fossil_label), hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(
    aes(label = label), size = 0.5, color = "grey30", offset = -0.25) +
  theme(legend.position = "bottom")
```

```{r fossil-map-plot-save, eval = params$doc_type == "doc"}
ggsave(
  plot = fossil_map_plot,
  file = (result_file(s_figure("fossil_map"), "pdf")),
  height = 70, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("fossil_map")`**.
Consensus phylogeny (Sanger dataset) with node labels indicating placement of fossil constraints (`r result_file(s_figure("fossil_map"), "pdf") %>% fs::path_file()`).
Number for each constraint is age in Ma.
Outgroup taxon *Zygnema circumcarinatum* (algae) pruned prior to dating analysis and not shown.

@@

```{r locus-upset-plot, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Make tibble with one row per locus per species
sanger_alignment_tbl %>%
  mutate(taxon = map(align_trimmed, rownames)) %>%
  select(locus = target, species = taxon) %>%
  unnest(cols = species) %>%
  # Exclude outgroups
  left_join(
    unique(select(sanger_sampling, species, outgroup)), by = "species") %>%
  filter(outgroup == FALSE) %>%
  select(-outgroup) %>%
  group_by(species) %>%
  summarize(locus = list(locus)) %>%
  ggplot(aes(x = locus)) +
  geom_bar() +
  labs(x = "Locus combination", y = "Count (num. species)") +
  scale_x_upset(n_intersections = 20)
```

@@

```{r bs-plot-make}
phy <- sanger_con_tree_rooted_pruned

@@

# Make plot
bs_plot <- 
  ggtree(
    phy,
    size = 0.07, color = "grey30") %<+% bs_tibble +
  # Add node points
  geom_nodepoint(
    aes(
      color = bs), size = 0.1, alpha = 0.9,
      shape = 16, fill = "transparent") +
  scale_color_viridis_c(
    "BS",
    option = "B", na.value = "transparent",
    labels = scales::percent
    ) +
  # Add tip labels
  geom_tiplab(
    aes(label = label), size = 0.5, color = "grey30", offset = -0.005) +
  theme(legend.position = "bottom")
```

```{r bs-plot-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bs_plot,
  file = result_file(s_figure("bs"), "pdf"),
  height = 70, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("bs")`**.
Consensus phylogeny (Sanger dataset) with node labels showing bootstrap support (%) (`r result_file(s_figure("bs"), "pdf") %>% fs::path_file()`).
Darker colors indicate lower support.
The backbone (plastome) phylogeny was used as a constraint tree, so all clades present in the backbone phylogeny tree receive 100% bootstrap support.
Outgroup taxon *Zygnema circumcarinatum* (algae) not shown.

@@

```{r backbone-full-plot-make}
# Don't show Zygnema in plot
phy <- ape::drop.tip(plastome_tree_rooted, "Zygnema_circumcarinatum")

@@

# Make plot
bb_full_plot <-
  ggtree(phy, size = 0.2, color = "grey10") %<+% 
  bs_tibble +
  # now we can show BS values using geom_text()
  geom_text(
    aes(label = bs), hjust = -.25, size = 1.0,
    color = "grey50") +
  # add tip labels
  geom_tiplab(aes(label = label), size = 1.0, offset = -0.005)
```

```{r backbone-full-plot-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bb_full_plot,
  file = (result_file(s_figure("backbone_full"), "pdf")),
  height = 20, width = 10, units = "in",
  limitsize = FALSE)
```

**`r s_figure("backbone_full")`**.
Consensus plastome (backbone) phylogeny (`r result_file(s_figure("backbone_full"), "pdf") %>% fs::path_file()`).
Values at nodes are bootstrap support (%); only values less than 100% are shown.
Outgroup taxon *Zygnema circumcarinatum* (algae) not shown.

@@

```{r du-ages-plot, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Helper function for subsetting dates to get into
# almost-long format
subset_ages <- function(other_dates, author_sel) {
  other_dates %>%
    select(family, contains(glue("{author_sel}_"))) %>%
    rename_with(~str_remove_all(., glue("{author_sel}_"))) %>%
    mutate(author = author_sel)
}

@@

```{r comp-fossils-plot, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
ts_fossil_calibration_tips %>%
  select(minimum_age, node_calibrated) %>%
  mutate(dataset = "ts") %>%
  bind_rows(
    con_fossil_calibration_tips %>%
      select(minimum_age, node_calibrated) %>%
      mutate(dataset = "current")
  ) %>%
  mutate(node_calibrated = fct_reorder(node_calibrated, minimum_age)) %>%
  ggplot(aes(x = minimum_age, y = node_calibrated, fill = dataset)) +
  geom_col(position = position_dodge()) +
  scale_x_continuous("Age (millions of years)", expand = expansion(mult = c(0,0.1))) +
  theme_bw() +
  scale_fill_manual(
    values = c(current = "grey30", ts = "grey80"),
    labels = c(current = "This Study", ts = "Testo and Sundue 2016")
  ) +
  theme(
    axis.title.y = element_blank(),
    legend.title = element_blank(),
    axis.text.y = element_text(size = 6),
    legend.position = "bottom"
  )
```

@@

```{r comp-stem-ages-plot, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
# Combine family stem ages into almost-long format:
# one row per family/study combination,
# cols for median, high, and low date
all_family_stem_ages <-
  bind_rows(
    subset_ages(other_dates, "ts"),
    transmute(
      ts_family_stem_ages,
      family,
      author = "current",
      median = age
    )
  ) %>%
  left_join(
    unique(select(sanger_sampling, family, major_clade)), by = "family") %>%
  left_join(clade_tibble, by = "major_clade") %>%
  # (some median dates are NA if that family wasn't included in a study)
  assert(not_na, family, major_clade, code, author) %>%
  # Set family as a factor ordered by current age estimate
  left_join(ts_family_stem_ages, by = "family") %>%
  mutate(family = fct_reorder(family, age)) %>%
  select(-age)

@@

```{r ts-fossil-map-plot-make}
# Make tibble with node number and label for ggtree
ts_node_labs_tbl <-
  ts_fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(
      taxon_1, taxon_2,
      ~ape::getMRCA(sanger_con_tree_rooted_pruned, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
ts_fossil_map_plot <-
  ggtree(
    sanger_con_tree_rooted_pruned,
    branch.length = "none",
    size = 0.1, color = "grey30") %<+% ts_node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 0.6) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"),
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(
    aes(label = fossil_label),
    hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(aes(label = label),
  size = 0.5, color = "grey30", offset = -0.25) +
  xlim(0, 66.1) +
  theme(legend.position = "bottom")
```

```{r ts-fossil-map-plot-save, eval = params$doc_type == "doc"}
ggsave(
  plot = ts_fossil_map_plot,
  file = (result_file(s_figure("ts_fossil_map"), "pdf")),
  height = 70, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("ts_fossil_map")`**.
Consensus phylogeny (Sanger dataset) with node labels indicating placement of fossil constraints of @Testo2016a.
Number for each constraint is age in Ma.
Outgroup taxon *Zygnema circumcarinatum* (algae) pruned prior to dating analysis and not shown.

@@

```{r comp-stem-points-plot, fig.width = 6, fig.height = 5, eval = params$doc_type != "manual"}
# For checking that model data is same as plot data
mod_fitted_data <-
ftol_ts_comp_mod %>%
  broom::augment() %>%
  arrange(ftol)

@@

```{r ltt-plot, fig.width = 7, fig.height = 6, eval = params$doc_type != "manual"}
# Make lineages through time (LTT) plot comparing Angiosperms and polypods
